---
title: "Forecasting Insufficient Demand in the Turkish Services Sector"
author: "Ömer Faruk ASLANER - 138722028"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
packages <- c("dplyr", "stringr", "lubridate", "ggplot2", "forecast", "tibble", "purrr", "tidyr", "knitr", "readr", "tuikr")
missing_packages <- packages[!vapply(packages, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_packages) > 0) install.packages(missing_packages)
library(dplyr); library(stringr); library(lubridate); library(ggplot2); library(forecast); library(tibble); library(purrr); library(tidyr); library(knitr); library(readr)
source("R/data_import.R")
source("R/accuracy_measures.R")
source("R/forecasting_methods.R")
source("R/plots.R")
dir.create("outputs/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("outputs/tables", recursive = TRUE, showWarnings = FALSE)
```

# Project Overview

This project forecasts the next quarterly value of **insufficient demand** as a factor limiting activities in the Turkish services sector. The selected TÜİK table is **Main factors limiting activities in services sector**, under **Short-Term Economic Indicators**.

# Student Information

- Student name: **Ömer Faruk ASLANER**
- Student number: **138722028**
- Selected variable: **Insufficient demand (%)**

# TÜİK Data Source and Connection

```{r data-import, message=FALSE, warning=FALSE}
tuik_result <- fetch_tuik_services_insufficient_demand()
series <- tuik_result$data
metadata <- tuik_result$metadata
```

```{r metadata-table, echo=FALSE}
metadata_table <- tibble::tibble(
  Field = c("Student name", "Student number", "TÜİK theme/category", "Theme ID", "TÜİK data set", "TÜİK table name", "tuikr dataflow ID", "Selected variable", "Frequency", "Unit", "Time coverage", "Number of observations", "Latest available observation", "Forecast target period", "Data access date", "Table URL"),
  Value = c(metadata$student_name, metadata$student_number, metadata$theme_category, metadata$theme_id, metadata$dataset_name, metadata$table_name, metadata$dataflow_id, metadata$selected_variable, metadata$data_frequency, metadata$unit, metadata$time_coverage, metadata$number_of_observations, metadata$latest_available_observation, metadata$forecast_target_period, metadata$data_access_date, metadata$table_url)
)
knitr::kable(metadata_table, caption = "TÜİK data source information")
```

**Source note:** `r metadata$source_note`

# Use of TÜİK Data in R

The selected table contains several possible limiting factors. In R, the table is reduced to the selected quarterly time series: **insufficient demand**. The period variable is constructed from the year and quarter values. The latest available observation is `r metadata$latest_available_observation`, and the next period to forecast is `r metadata$forecast_target_period`.

```{r data-preview, echo=FALSE}
series |>
  dplyr::select(obsTime, year, quarter, value) |>
  dplyr::rename(`Insufficient demand (%)` = value) |>
  head(12) |>
  knitr::kable(caption = "First observations of the selected series")
```

# Exploratory Time Series Analysis

```{r descriptive-statistics, echo=FALSE}
series |>
  dplyr::summarise(
    Observations = n(),
    Minimum = min(value),
    Maximum = max(value),
    Mean = mean(value),
    Median = median(value),
    `Standard deviation` = sd(value)
  ) |>
  knitr::kable(digits = 3, caption = "Descriptive statistics")
```

```{r actual-series-plot, fig.width=8, fig.height=4.5, echo=FALSE}
p_actual <- plot_series(series)
print(p_actual)
ggplot2::ggsave("outputs/figures/actual_series_plot.png", p_actual, width = 8, height = 4.5)
```

The series is quarterly, so seasonal forecasting methods are applicable. The COVID-19 period also introduces visible volatility, which is considered as a limitation.

# Forecasting Methods Applied

```{r forecasting, message=FALSE, warning=FALSE}
forecast_results <- build_candidate_forecasts(series)
methods <- forecast_results$methods
accuracy <- forecast_results$accuracy
best_method <- accuracy$Method[1]
best_forecast <- accuracy$Next_Period_Forecast[1]
readr::write_csv(accuracy, "outputs/tables/accuracy_comparison.csv")
final_forecast <- tibble::tibble(
  Selected_Superior_Method = best_method,
  Data_Access_Date = metadata$data_access_date,
  Latest_Available_Observation = metadata$latest_available_observation,
  Forecast_Target_Period = metadata$forecast_target_period,
  Forecasted_Value = best_forecast,
  Bias = accuracy$Bias[1],
  MAD = accuracy$MAD[1],
  MSE = accuracy$MSE[1],
  MAPE = accuracy$MAPE[1],
  RSFE = accuracy$RSFE[1],
  Tracking_Signal = accuracy$Tracking_Signal[1]
)
readr::write_csv(final_forecast, "outputs/tables/final_forecast.csv")
save_all_method_plots(series, methods, best_method)
```

The required methods are all considered because the selected series is a positive quarterly time series with sufficient observations for seasonal methods and decomposition.

# Forecast Accuracy Comparison

```{r accuracy-table, echo=FALSE}
accuracy |>
  dplyr::select(Method, Bias, MAD, MSE, MAPE, RSFE, Tracking_Signal, Next_Period_Forecast, Note) |>
  knitr::kable(digits = 3, caption = "Forecast accuracy comparison")
```

# Actual vs Forecast Plots

```{r method-plots, fig.width=8, fig.height=4.5, echo=FALSE}
for (method_name in names(methods)) {
  print(plot_method_fit(series, methods, method_name))
}
```

# Superior Method Selection

The superior method is selected as **`r best_method`** because it provides the strongest accuracy performance in the comparison table while remaining suitable for a quarterly series. The final decision considers both numerical error measures and the visual fit between actual and forecasted values.

# Final Next-Period Forecast

```{r final-forecast-table, echo=FALSE}
final_forecast |>
  knitr::kable(digits = 3, caption = "Final forecast result")
```

The forecast for **`r metadata$forecast_target_period`** is **`r round(best_forecast, 3)`%**. This value represents the expected percentage of service-sector enterprises that may report insufficient demand as a factor limiting activities in the next quarter.

# Interpretation of Results

The selected variable is meaningful because insufficient demand reflects demand-side pressure in the services sector. A higher value suggests that more enterprises experience insufficient market demand, while a lower value suggests a relatively stronger business environment. The quarterly structure makes seasonal methods relevant, and the comparison table shows which method best captures the historical pattern.

# Limitations

The analysis is limited by structural shocks, especially the COVID-19 period, possible survey methodology changes, and the fact that the model uses only historical values of the selected variable. No external explanatory variables are included.

# Reproducibility

To reproduce the project, open the repository in RStudio and run:

```r
source("run_all.R")
```

This renders `forecasting_project.html` and writes forecast tables and figures under `outputs/`.
